Improving the Accuracy of a Sequential Mining Model for Hurricane Trajectory Prediction

### 2018-01-29 paper/manuscript_revised_jan25.tex
Slightly shorten AprioriAll section. Reword weighted data contribution

The hurricane trajectory model in this work does not implement step (1), as the training data from HURDAT2 was instead translated into a $\langle HurricaneID, CoordinateSequence\rangle$ database as a preprocessing step before any aspects of the model are executed. Figure (1) shows an example of how the trajectory database is structured after the initial mapping stage. Only two dimensions are needed in the database, being the hurricane ID and the sequence of latitude-longitude coordinates. All hurricanes in HURDAT2 have an ID corresponding to the location (i.e. the Atlantic), the index ordered by appearance in the hurricane year, and the year itself. For example, hurricane Katrina of year 2005 corresponds to AL122005. By directly translating the data instead of processing it with AprioriAll, all trajectories in the original database are preserved. This includes coordinate sequences which may ultimately be infrequent. The sequential mapping of AprioriAll is performed as a preprocessing step so that the database used in the prediction model will always consist of hurricanes and their coordinate sequences. The prepared (mapped) database may therefore be mined directly without the need for repeated mapping in subsequent executions of AprioriAll.

@@

Assuming the input data is ordered chronologically, introducing weighted support to AprioriAll will result in higher support for patterns which occur in recent years. Recent trajectories will require less repetition to produce association rules, and higher repetition will be needed to generate frequent sequences from older data. The approach of chronologically weighted data will produce association rules that favour recent trends in hurricane trajectories. Therefore, the introduction of weighted support to AprioriAll is expected to further increase the hurricane trajectory prediction correctness ratio for current and future trajectories being tested upon. Publications such as \cite{war-mining} and \cite{cai-weighted-thesis} first introduce the use of weighted support for association rule mining. However, to the knowledge of the authors this work is the first to (1) apply weighted support to sequential data, (2) weigh individual records based on a proportional scale and (3) allow records to have a weight less than 1.

>>>>

The hurricane trajectory model in this work does not implement step (1), as the training data from HURDAT2 was instead translated into a $\langle HurricaneID, CoordinateSequence\rangle$ database as a preprocessing step before launching the model. Figure (1) shows an example of how the trajectory database is structured after the initial mapping stage. Only two dimensions are needed in the database, being the hurricane ID and the sequence of latitude-longitude coordinates. All hurricanes in HURDAT2 have an ID corresponding to the location (i.e. the Atlantic), the index ordered by appearance in the hurricane year, and the year itself. For example, hurricane Katrina of year 2005 corresponds to AL122005. By directly translating the data instead of processing it with AprioriAll, all trajectories in the original database are preserved. This includes coordinate sequences which may ultimately be infrequent. The sequential mapping of AprioriAll is performed as a preprocessing step so that the database used in the prediction model will always consist of hurricanes and their coordinate sequences, removing the need for repeated mapping in subsequent executions of AprioriAll.

@@

Assuming the input data is ordered chronologically, introducing weighted support to AprioriAll will result in higher support for patterns which occur in recent years. Recent trajectories will require less repetition to produce association rules, and higher repetition will be needed to generate frequent sequences from older data. The approach of chronologically weighted data will produce association rules that favour recent trends in hurricane trajectories. Therefore, the introduction of weighted support to AprioriAll is expected to further increase the hurricane trajectory prediction correctness ratio for current and future trajectories being tested upon. Publications such as \cite{war-mining} and \cite{cai-weighted-thesis} first introduce the use of weighted support for association rule mining. However, to the knowledge of the authors this work is the first to (1) apply weighted support to sequential data, (2) apply weight to records as opposed to individual items and (3) allow records to have a weight less than 1.
